---
title: "Covid91 vaccine study Final2024"
author: "Eliana Martinez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effetive in males?Add commentMore actions

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectivness on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectivness on male.

cat-cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```

## Graphical ResultsAdd commentMore actions

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```

This graph demonstrates that the vaccine is effective in preventing COVID-19 in males, as fewer men who received the vaccine contracted the virus compared to those who received the placebo.

## Numerical ResultsAdd commentMore actions


```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The table shows that 26.20% of men who received the vaccine contracted COVID-19, while 73.80% of men who received the placebo became infected.


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The analysis examined the relationship between COVID-19 infection status and treatment type (vaccine vs. placebo) among males in the dataset. Results showed a significantly higher infection rate in the placebo group (2.82%) compared to the vaccine group (1.02%). A chi-square test confirmed a strong association between treatment and infection status (χ² = 77.87, *p* < 0.0001), and a Fisher's exact test supported this finding with an odds ratio of approximately 2.82, indicating that males in the placebo group were nearly three times more likely to be infected than those who received the vaccine. These results suggest that the vaccine was effective in reducing COVID-19 infections among males in the study.

```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

The p-value from Fisher’s test is effectively zero, which is far below the 0.05 threshold. Therefore, we reject the null hypothesis, indicating that the vaccine is effective in males. The odds ratio is 2.81, meaning that men who received the vaccine were nearly three times less likely to contract COVID-19 compared to those who received the placebo.Add commentMore actions

Furthermore, the 95% confidence interval ranges from 2.21 to 3.61. Since the entire interval is above 1, this provides additional evidence supporting the vaccine’s effectiveness. Overall, the data strongly suggest that the vaccine is a beneficial option for males.Add commentMore actions


## Females
##Methods  Add commentMore actions
$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat.

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical ResultsAdd commentMore actions

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph shows that the vaccine is more effective in women, as fewer women who received the vaccine contracted COVID-19 compared to those who received the placebo.

##Inferential Results Add commentMore actions

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.00000001914, which is less than 0.05, so we reject the null hypothesis. The odds ratio from Fisher’s exact test is 1.856, indicating that the vaccine is approximately 1.86 times more effective in women who received it compared to those who received the placebo.

```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`
Fisher’s test yielded a p-value of 1e-08, which is significantly lower than the 0.05 significance level, leading us to reject the null hypothesis. This indicates that the vaccine is also effective in females. The odds ratio of 1.9 suggests that vaccinated women were nearly twice as likely to avoid contracting COVID-19 compared to those who received a placebo.

The 95% confidence interval ranges from 1.5 to 2.3, entirely above 1, reinforcing the conclusion that the vaccine has a statistically significant impact. These findings support the vaccine as a beneficial option for women.


## LGBTQ

##Methods  Add commentMore actions
$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ. 

Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```

##Graphical Results  Add commentMore actions

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph indicates that the vaccine is effective in preventing COVID-19 among individuals who identify as LGBTQ.

##Numerical Results Add commentMore actions

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The data indicates that only 20% of vaccinated LGBTQ individuals contracted COVID-19, in contrast to 80% of those who received a placebo. This implies that the vaccine effectively reduces COVID-19 infections among members of the LGBTQ community.

##Infrential ResultsAdd commentMore actions

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The test produced a p-value of 0.000002266, which is well below the 0.05 threshold, leading to the rejection of the null hypothesis. According to Fisher’s exact test, the odds ratio is 4.27, meaning that vaccinated individuals were over four times more likely to avoid contracting COVID-19 than those who received the placebo.

## Druggies

##Methods  Add commentMore actions

$H_0$ The preventative vaccine does not exhibit its effective in people who are drug users.    
$H_A$ The preventative vaccine exhibits its effective in people who are drug users.  

Cat~Cat

```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
```

##Graphical ResultsAdd commentMore actions

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The bar graph shows that the vaccine appears to be less effective in individuals who use drugs.

##Numerical Results Add commentMore actions

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The table shows that 91.67% of drug users who contracted COVID-19 had also received the vaccine. This suggests that the vaccine may be less effective among individuals who use drugs.

##Inferential Results    Add commentMore actions

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.0000000001414, which is less than 0.05, so we reject the null hypothesis. However, despite the statistical significance, the results suggest that the vaccine may have limited effectiveness in this particular group.

```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`Add commentMore actions

The p-value is 1e-08, which is extremely low and well below the 0.05 threshold, so we reject the null hypothesis. This indicates that the odds of contracting COVID-19 differ significantly depending on whether or not a person received the vaccine.


# Overall Results and Conclusions

Men:
The data suggests the vaccine is effective for men. Only 26.2% of vaccinated men contracted COVID-19, compared to 73.8% of those who received the placebo—a notable difference. The p-value was 2.2e-16, far below the 0.05 threshold, allowing us to confidently reject the null hypothesis. The odds ratio was approximately 2.8, indicating that vaccinated men were about 2.8 times more likely to avoid infection than those who received the placebo. These results suggest the vaccine is a reliable preventive measure for men.

Women:
Similarly, the vaccine appears effective for women. Among those who contracted COVID-19, 35.07% had been vaccinated, while 64.93% had received the placebo. The p-value was 1.914e-8, well under the 0.05 threshold, supporting rejection of the null hypothesis. The odds ratio was about 1.86, showing that vaccinated women were significantly less likely to contract COVID-19. These findings indicate that the vaccine helps reduce infection risk among females.

LGBTQ Individuals:
For LGBTQ individuals, the vaccine also shows strong effectiveness. Just 20% of COVID-19 cases in this group occurred among vaccinated individuals, compared to 80% among those who received a placebo. This substantial gap is supported by a p-value of 2.266e-6, well below the 0.05 cutoff, leading to rejection of the null hypothesis. The odds ratio of 4.27 suggests that vaccinated LGBTQ individuals were over four times more likely to avoid infection. These results provide strong evidence of the vaccine’s benefit for this group.

Drug Users:
In contrast, the vaccine appears less effective among individuals who use drugs. Surprisingly, 91.67% of those in this group who contracted COVID-19 had received the vaccine, while only 8.33% had received the placebo—opposite of the expected pattern. The p-value was 1.414e-10, well below 0.05, indicating statistical significance. However, the odds ratio was just 0.087, suggesting a reduced likelihood of protection in this group. This result raises concerns about the vaccine’s effectiveness among drug users and may warrant further investigation.
